\bf Classification and Regression on Random Dot Product Graphs

### 2024-03-22 draft.Rmd
added another data analysis

@@

@@

In the first example, we analyzed fiber count data between brain regions from the Human Connectome Project (HCP). 
When analyzing these data as graphs, we denote the regions as vertices and the fiber counts between pairs of regions as weighted edges. 
A plausible statistical model for these data is to assume that the edge weights between pairs of vertices is Poisson distributed, i.e., the adjacency matrix is sampled as $A_{ij} \indep \Poisson(\Theta_{ij})$, where $\Theta \in \mathbb{R}_+^{n \times n}$ is a symmetric matrix of Poisson parameters.

@@

dplyr::tibble(
  Metric = c('Angle between hemispheres',
             'Edge connectivity within left hemisphere',
             'Edge connectivity within right hemisphere',
             'Edge connectivity between hemispheres',
             'Degree within hemisphere',
             'Degree between hemispheres',
             'Number of triangles in left hemisphere',
             'Number of triangles in right hemisphere',
             'Assortativity w.r.t. hemisphere',
             # 'Transitivity',
             'Modularity w.r.t. hemisphere'),
  Correlation = c(cor(ase.stats.df$age, ase.stats.df$angle.diff),
                  cor(ase.stats.df$age, ase.stats.df$B.ll),
                  cor(ase.stats.df$age, ase.stats.df$B.rr),
                  cor(ase.stats.df$age, ase.stats.df$B.lr),
                  cor(ase.stats.df$age, ase.stats.df$deg.within),
                  cor(ase.stats.df$age, ase.stats.df$deg.between),
                  cor(assort.df$age, assort.df$n_tri1),
                  cor(assort.df$age, assort.df$n_tri2),
                  cor(assort.df$age, assort.df$assortativity),
                  # cor(assort.df$age, assort.df$transitivity),
                  cor(assort.df$age, assort.df$modularity)),
  `95% conf. int.` = c(cor.test(ase.stats.df$age, ase.stats.df$angle.diff)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(ase.stats.df$age, ase.stats.df$B.ll)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(ase.stats.df$age, ase.stats.df$B.rr)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(ase.stats.df$age, ase.stats.df$B.lr)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(ase.stats.df$age, ase.stats.df$deg.within)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(ase.stats.df$age, ase.stats.df$deg.between)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(assort.df$age, assort.df$n_tri1)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(assort.df$age, assort.df$n_tri2)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(assort.df$age, assort.df$assortativity)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       # cor.test(assort.df$age, assort.df$transitivity)$conf.int %>% 
                       #   round(3) %>% 
                       #   paste(collapse = ', ') %>% 
                       #   paste0('(', ., ')'),
                       cor.test(assort.df$age, assort.df$modularity)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'))) %>% 
  knitr::kable(digits = 3, caption = 'Correlation between age and various graph metrics')
```

>>>>

In the first regression simulation,

In the next regression simulation, we applied

@@

## Drosophila Connectome

In the second example, we analyzed the larval *Drosophila* mushroom body connectome \citep{Eichler141762}, which has been studied as a GRDPG by \citet{athreya2020estimation}. 
This dataset consists of two graphs representing two networks of neurons, one for each hemisphere of the *Drosophila* brain. 
In these graphs, each vertex is a neuron, and the labels correspond to one of four neuron types (Kenyon Cells, Input Neurons, Output Neurons, and Projection Neurons). 
The resulting graphs are illustrated in figure \ref{fig:mbconnectome-graph}.

```{r mbconnectome-graph, fig.cap = 'Graphs of the Drosophila connectomes. The left and right are of the left and right hemispheres, respectively. Each vertex represents a neuron, which are labeled by neuron type.', fig.height = 2, figh.width = 12, cache = TRUE}
library(mbstructure)
data(MBconnectome)

graph.r <- generate.graph(newrdat, vdf.right)
A.r <- as.matrix(igraph::as_adjacency_matrix(graph.r$g, type = 'both'))
z.r <- as.numeric(graph.r$vdf$type)
n.r <- length(z.r)
A.r <- sign(A.r + t(A.r))
diag(A.r) <- 0
graph.r <- ggraph(A.r, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n.r) / 2) + 
  geom_node_point(aes(colour = factor(z.r)), 
                  show.legend = FALSE) + 
  labs(colour = NULL) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

graph.l <- generate.graph(newldat, vdf.left)
A.l <- as.matrix(igraph::as_adjacency_matrix(graph.l$g, type = 'both'))
z.l <- as.numeric(graph.l$vdf$type)
n.l <- length(z.l)
A.l <- sign(A.l + t(A.l))
diag(A.l) <- 0
graph.l <- ggraph(A.l, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n.l) / 2) + 
  geom_node_point(aes(colour = factor(z.l)), 
                  show.legend = FALSE) + 
  labs(colour = NULL) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(graph.l, graph.r, nrow = 1)
```

```{r mbconnectome-ase, fig.cap = 'ASEs of the Drosophila connectome graphs. The embedding vectors are labeled by neuron type. ', fig.height = 2, fig.width = 8}
Xhat.l <- embedding(A.l, 2, 1)
ase.l <- ggplot() + 
  geom_point(aes(x = Xhat.l[, 1], y = Xhat.l[, 2],
                 colour = factor(z.l))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL)

Xhat.r <- embedding(A.r, 2, 1)
ase.r <- ggplot() + 
  geom_point(aes(x = Xhat.r[, 1], y = Xhat.r[, 2],
                 colour = factor(z.r))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL)

gridExtra::grid.arrange(ase.l, ase.r, nrow = 1)
```

When analyzed as a GRDPG, the ASE of each hemisphere suggests that nodes of each type falls along a curve in the latent space (figure \ref{fig:mbconnectome-ase}). 
This matches observations by \citet{athreya2020estimation}. 
In our analysis, we set the embedding dimension to $d = 4$, with 3 assortative dimensions and 1 disassortative dimension. 
We fit three latent structure Bezier curves, one for Kenyon Cells, one for Input and Output Neurons, and one for Projection Neurons, to the embedding for each hemisphere.
Then we fit a Beta distribution to to the timepoints along each curve and extracted the two Beta parameter estimates (via likelihood maximization) for each curve. 
If these Beta parameters are informative, we would expect the parameters for each hemisphere to match by neuron type, which is what we observe in these data (fig \ref{fig:mbconnectome-beta}).

```{r mbconnectome-beta, fig.cap = 'Beta parameter estimates for each curve and hemisphere.', fig.height = 3, fig.width = 6}
param.est.df <- readr::read_csv('~/dev/multilayer-rdpg/drosophila.csv')

param.est.df %>% 
ggplot() + 
  geom_point(aes(x = shape1, y = shape2, colour = region, shape = hemisphere)) +
  labs(x = expression(hat(alpha)),
       y = expression(hat(beta)),
       colour = 'Neuron type',
       shape = 'Hemisphere') +
  coord_fixed() + 
  theme_bw()
```

@@

In the second example, we analyzed fiber count data between brain regions from the Human Connectome Project (HCP). 
When analyzing these data as graphs, we denote the regions as vertices and the fiber counts between pairs of regions as weighted edges. 
A plausible statistical model for these data is to assume that the edge weights between pairs of vertices is Poisson distributed, i.e., the adjacency matrix is sampled as $A_{ij} \indep \Poisson(\Theta_{ij})$, where $\Theta \in \mathbb{R}_+^{n \times n}$ is a symmetric matrix of Poisson parameters.

@@

# https://github.com/jesusdaniel/JEG
source('~/dev/JEG/Code/joint_embedding.R')
A.list <- lapply(ids, function(id) {
  A <- dat.NumFibers[[id]] %>% 
    as.matrix()
  A <- (A + t(A)) / 2
  diag(A) <- 0
  return(A)
})
jeg.out <- multidembed(A.list, 2)
jeg.lm <- lm(ase.stats.df$age ~ jeg.out$lambda[, 1] + jeg.out$lambda[, 2])
jeg.yhat <- jeg.lm$fitted.values

dplyr::tibble(
  Metric = c('Angle between hemispheres',
             'Edge connectivity within left hemisphere',
             'Edge connectivity within right hemisphere',
             'Edge connectivity between hemispheres',
             'Degree within hemisphere',
             'Degree between hemispheres',
             'Number of triangles in left hemisphere',
             'Number of triangles in right hemisphere',
             'Assortativity w.r.t. hemisphere',
             # 'Transitivity',
             'Modularity w.r.t. hemisphere',
             'Joint Embedding'),
  Correlation = c(cor(ase.stats.df$age, ase.stats.df$angle.diff),
                  cor(ase.stats.df$age, ase.stats.df$B.ll),
                  cor(ase.stats.df$age, ase.stats.df$B.rr),
                  cor(ase.stats.df$age, ase.stats.df$B.lr),
                  cor(ase.stats.df$age, ase.stats.df$deg.within),
                  cor(ase.stats.df$age, ase.stats.df$deg.between),
                  cor(assort.df$age, assort.df$n_tri1),
                  cor(assort.df$age, assort.df$n_tri2),
                  cor(assort.df$age, assort.df$assortativity),
                  # cor(assort.df$age, assort.df$transitivity),
                  cor(assort.df$age, assort.df$modularity),
                  cor(assort.df$age, jeg.yhat)),
  `95% conf. int.` = c(cor.test(ase.stats.df$age, ase.stats.df$angle.diff)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(ase.stats.df$age, ase.stats.df$B.ll)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(ase.stats.df$age, ase.stats.df$B.rr)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(ase.stats.df$age, ase.stats.df$B.lr)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(ase.stats.df$age, ase.stats.df$deg.within)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(ase.stats.df$age, ase.stats.df$deg.between)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(assort.df$age, assort.df$n_tri1)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(assort.df$age, assort.df$n_tri2)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(assort.df$age, assort.df$assortativity)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       # cor.test(assort.df$age, assort.df$transitivity)$conf.int %>% 
                       #   round(3) %>% 
                       #   paste(collapse = ', ') %>% 
                       #   paste0('(', ., ')'),
                       cor.test(assort.df$age, assort.df$modularity)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'),
                       cor.test(assort.df$age, jeg.yhat)$conf.int %>% 
                         round(3) %>% 
                         paste(collapse = ', ') %>% 
                         paste0('(', ., ')'))) %>% 
  knitr::kable(digits = 3, caption = 'Correlation between age and various graph metrics')

### 2024-03-22 draft.tex
added another data analysis

@@

In the first example, we analyzed fiber count data between brain regions
from the Human Connectome Project (HCP). When analyzing these data as
graphs, we denote the regions as vertices and the fiber counts between
pairs of regions as weighted edges. A plausible statistical model for
these data is to assume that the edge weights between pairs of vertices
is Poisson distributed, i.e., the adjacency matrix is sampled as
\(A_{ij} \indep \Poisson(\Theta_{ij})\), where
\(\Theta \in \mathbb{R}_+^{n \times n}\) is a symmetric matrix of
Poisson parameters.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5915}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1690}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2394}}@{}}
\caption{Correlation between age and various graph
metrics}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Metric
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Correlation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
95\% conf. int.
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Metric
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Correlation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
95\% conf. int.
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Angle between hemispheres & 0.558 & (0.496, 0.615) \\
Edge connectivity within left hemisphere & 0.241 & (0.158, 0.321) \\
Edge connectivity within right hemisphere & 0.243 & (0.16, 0.322) \\
Edge connectivity between hemispheres & -0.436 & (-0.503, -0.363) \\
Degree within hemisphere & 0.436 & (0.363, 0.503) \\
Degree between hemispheres & -0.499 & (-0.561, -0.431) \\
Number of triangles in left hemisphere & -0.025 & (-0.111, 0.062) \\
Number of triangles in right hemisphere & 0.002 & (-0.084, 0.088) \\
Assortativity w.r.t. hemisphere & 0.436 & (0.363, 0.503) \\
Modularity w.r.t. hemisphere & 0.434 & (0.362, 0.502) \\
\end{longtable}

>>>>

\hypertarget{drosophila-connectome}{%
\subsection{Drosophila Connectome}\label{drosophila-connectome}}

In the second example, we analyzed the larval \emph{Drosophila} mushroom
body connectome \citep{Eichler141762}, which has been studied as a GRDPG
by \citet{athreya2020estimation}. This dataset consists of two graphs
representing two networks of neurons, one for each hemisphere of the
\emph{Drosophila} brain. In these graphs, each vertex is a neuron, and
the labels correspond to one of four neuron types (Kenyon Cells, Input
Neurons, Output Neurons, and Projection Neurons). The resulting graphs
are illustrated in figure \ref{fig:mbconnectome-graph}.

{\centering \includegraphics{draft_files/figure-latex/mbconnectome-graph-1}

\caption{Graphs of the Drosophila connectomes. The left and right are of the left and right hemispheres, respectively. Each vertex represents a neuron, which are labeled by neuron type.}\label{fig:mbconnectome-graph}
\end{figure}

{\centering \includegraphics{draft_files/figure-latex/mbconnectome-ase-1}

\caption{ASEs of the Drosophila connectome graphs. The embedding vectors are labeled by neuron type. }\label{fig:mbconnectome-ase}
\end{figure}

When analyzed as a GRDPG, the ASE of each hemisphere suggests that nodes
of each type falls along a curve in the latent space (figure
\ref{fig:mbconnectome-ase}). This matches observations by
\citet{athreya2020estimation}. In our analysis, we set the embedding
dimension to \(d = 4\), with 3 assortative dimensions and 1
disassortative dimension. We fit three latent structure Bezier curves,
one for Kenyon Cells, one for Input and Output Neurons, and one for
Projection Neurons, to the embedding for each hemisphere. Then we fit a
Beta distribution to to the timepoints along each curve and extracted
the two Beta parameter estimates (via likelihood maximization) for each
curve. If these Beta parameters are informative, we would expect the
parameters for each hemisphere to match by neuron type, which is what we
observe in these data (fig \ref{fig:mbconnectome-beta}).

{\centering \includegraphics{draft_files/figure-latex/mbconnectome-beta-1}

\caption{Beta parameter estimates for each curve and hemisphere.}\label{fig:mbconnectome-beta}
\end{figure}

@@

In the second example, we analyzed fiber count data between brain
regions from the Human Connectome Project (HCP). When analyzing these
data as graphs, we denote the regions as vertices and the fiber counts
between pairs of regions as weighted edges. A plausible statistical
model for these data is to assume that the edge weights between pairs of
vertices is Poisson distributed, i.e., the adjacency matrix is sampled
as \(A_{ij} \indep \Poisson(\Theta_{ij})\), where
\(\Theta \in \mathbb{R}_+^{n \times n}\) is a symmetric matrix of
Poisson parameters.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5915}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1690}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2394}}@{}}
\caption{Correlation between age and various graph
metrics}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Metric
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Correlation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
95\% conf. int.
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Metric
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Correlation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
95\% conf. int.
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Angle between hemispheres & 0.558 & (0.496, 0.615) \\
Edge connectivity within left hemisphere & 0.241 & (0.158, 0.321) \\
Edge connectivity within right hemisphere & 0.243 & (0.16, 0.322) \\
Edge connectivity between hemispheres & -0.436 & (-0.503, -0.363) \\
Degree within hemisphere & 0.436 & (0.363, 0.503) \\
Degree between hemispheres & -0.499 & (-0.561, -0.431) \\
Number of triangles in left hemisphere & -0.025 & (-0.111, 0.062) \\
Number of triangles in right hemisphere & 0.002 & (-0.084, 0.088) \\
Assortativity w.r.t. hemisphere & 0.436 & (0.363, 0.503) \\
Modularity w.r.t. hemisphere & 0.434 & (0.362, 0.502) \\
Joint Embedding & 0.556 & (0.493, 0.613) \\
\end{longtable}
